---
title: "Class Exercise - Weather Forecasting"
author: "Steve Spence"
date: "3/3/2020"
output: word_document
---

In class exercise on March 3, 2020 for weather forecasting.

Use any or all of the methods to improve weather forecasting for the problem we discussed in class. For example, by adjusting the number of units in each recurrent layer in the stacked setup, or using layer_lstm() instead of layer_gru().

You can also try experimenting with a combination of 1d_convnets and rnn.

Don't forget to eventually run the best-performing models (in terms of validation MAE) on the test set!

Download and uncompress the data as follows and take a glimpse at the dataset

```{r, eval=FALSE, echo=TRUE}

library(tibble)
library(readr)

dir.create("~/Downloads/jena_climate", recursive = TRUE)
download.file(
  "https://s3.amazonaws.com/keras-datasets/jena_climate_2009_2016.csv.zip",
  "~/Downloads/jena_climate/jena_climate_2009_2016.csv.zip"
)
unzip(
  "~/Downloads/jena_climate/jena_climate_2009_2016.csv.zip",
  exdir = "~/Downloads/jena_climate"
)

data_dir <- "~/Downloads/jena_climate"
fname <- file.path(data_dir, "jena_climate_2009_2016.csv")
data <- read_csv(fname)

glimpse(data)

```

Reviewing a plot of the data:

```{r}
library(ggplot2)
ggplot(data, aes(x = 1:nrow(data), y = `T (degC)`)) + geom_line()
```

## Preparing the Data

We will use the following parameter values:

* `lookback = 1440`, i.e. our observations will go back 10 days.
* `steps = 6`, i.e. our observations will be sampled at one data point per hour.
* `delay = 144`, i.e. our targets will be 24 hours in the future.

First, you'll convert the R data frame which we read earlier into a matrix of floating point values (we'll discard the first column which included a text timestamp):

```{r}
data <- data.matrix(data[,-1])
```

You'll then preprocess the data by subtracting the mean of each timeseries and dividing by the standard deviation. You're going to use the first 200,000 timesteps as training data, so compute the mean and standard deviation for normalization only on this fraction of the data.

```{r}
train_data <- data[1:200000,]
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)
data <- scale(data, center = mean, scale = std)
```

```{r}
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index))
    max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size-1, max_index))
      i <<- i + length(rows)
    }
    
    samples <- array(0, dim = c(length(rows), 
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
                     
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]] - 1, 
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,2]
    }            
    
    list(samples, targets)
  }
}
```


```{r}
lookback <- 1440
step <- 6
delay <- 144
batch_size <- 128

train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 200000,
  shuffle = TRUE,
  step = step, 
  batch_size = batch_size
)

val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 200001,
  max_index = 300000,
  step = step,
  batch_size = batch_size
)

test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 300001,
  max_index = NULL,
  step = step,
  batch_size = batch_size
)

# This is how many steps to draw from `val_gen`
# in order to see the whole validation set:
val_steps <- (300000 - 200001 - lookback) / batch_size

  # This is how many steps to draw from `test_gen`
# in order to see the whole test set:
test_steps <- (nrow(data) - 300001 - lookback) / batch_size
```

```{r}
evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],2]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}
```


```{r, echo=TRUE, results='hide'}
library(keras)

model <- keras_model_sequential() %>% 
  layer_flatten(input_shape = c(lookback / step, dim(data)[-1])) %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

```{r}

plot(history)

```


## Baseline Recurrent Model

```{r echo=TRUE, results='hide'}
model <- keras_model_sequential() %>% 
  layer_gru(units = 32, input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

Results:

```{r}
plot(history)
```


## Improvements to Model

```{r, echo=TRUE, results='hide'}

model <- keras_model_sequential() %>% 
  layer_gru(units = 32, 
            dropout = 0.2,                                      # Recurrent dropout set to 0.2
            recurrent_dropout = 0.6,
            return_sequences = TRUE,                            # Recurrent dropout set to 0.6
            input_shape = list(NULL, dim(data)[[-1]])) %>%
  layer_gru(units = 64,                                         # Added additional gru layer 64 units
            activation = "relu",                                # Relu activation
            dropout = 0.2,                                      # Recurrent dropout set to 0.2
            recurrent_dropout = 0.6) %>%                        # Recurrent dropout set to 0.6
  layer_dense(units = 32, activation = "relu") %>%              # Added additional dense layer 32 units
  layer_dense(units = 64, activation = "relu") %>%              # Added additional dense layer 64 units
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 40,
  validation_data = val_gen,
  validation_steps = val_steps
)

```

```{r}
plot(history)
```

Review of test results:

```{r}

history_test <- model %>% fit_generator(
  test_gen,
  steps_per_epoch = 500,
  epochs = 40,
  test_data = test_gen,
  test_steps = test_steps
)


```





